Draft TLF generator

### 2023-04-26 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

---
title: 'Draft TLF generator'
date: '2023-04-25'
author: Qing Zou
categories: ['R', 'R6class']
description: 'Draft the TLF generator tool'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

@@

```{r rcell15}
#| echo: true
#| eval: true
#' @title AdsModel Class
#'
#'
#' @description
#' This is the  base class for ADaM dataset, to manage the CDISC attributes on dataset level
#'
#' AdsModel are build around the three following key parts:
#'
#' * Methods
#'
#'
#' For now, only implement that adding the attribute Model label by default. see `add_model_name()`
#'
#' test <- AdsModel$new()
#' test$initialize(adsl, adpsl)
#'
#' names(test$dataset)
#' a <- test$dataset
#'
#' tmp <- test$get_dataset('adsl')
#' attributes(tmp)
#'
#' adsl2 <- adsl
#' test$add_dataset(adsl2)
#' names(test$dataset)
#'
#' ```
#'

AdsModel <- R6::R6Class(
  "AdsModel",
  public = list(
    lib = NA,
    initialize = function(...){
      df_lst <-  tibble::lst(...)
      if(is.character(check_list(df_lst, min.len = 1))){
        # message("No datasets passed")
      }else{
        df_lst <- purrr::map2(df_lst, names(df_lst), ~utils$add_model_name(.x, .y))
        private$.dataset <- private$.dataset |> append(df_lst)
      }

    },
    get_dataset = function(domain){
      if(missing(domain)){
        return(private$.dataset)
      }else{
        checkmate::check_character(domain,
                                   min.chars = 1L,
                                   any.missing = FALSE)
        if(length(domain) == 1){
          return(purrr::keep_at(private$.dataset, domain)[[1]])
        }else{
          return(purrr::keep_at(private$.dataset, domain))
        }

@@

      },
    load_dataset = function(path = NULL, domains){

@@

```{r}
#| echo: true
#| eval: true
# test <- AdsModel$new(adsl)
adsl <- haven::read_xpt('../../data/adam/adsl.xpt')
test <- AdsModel$new()

@@

names(test$dataset)
```

@@

tmp <- test$get_dataset('adsl')
attributes(tmp)
```

@@

adtte <- haven::read_xpt('../../data/adam/adtte.xpt')
test$add_dataset(adtte)
names(test$dataset)
tmp2 <- test$get_dataset('adtte')
attributes(tmp2)
```

@@

VariableProc <- R6::R6Class(
  "VariableProc",
  inherit = AdsModel,
  public = list(
    # datasets level dictionary, can be retrieved by dataset
    # variables level dictionary, can be retrieved by dataset$variable
    dictionary = list(),
    initialize = function(dictionary_path = NA){

@@

@@

attributes(adsl_tmp)

attributes(adsl_tmp$SEX)$format

@@

```{r}
test2$get_decode()

adsl_tmp2 <- test2$get_dataset('adsl')
attributes(adsl_tmp)
attributes(adsl_tmp2$SEX)$format

@@

```{r}
TLFProc <- R6::R6Class(
  "TLFProc",
  public = list(
    tlfs = NULL,
    tlfmetadata = NULL,
    initialize = function(config = NULL) {
      self$tlfs = new.env(parent = emptyenv())

      if(!is.null(config)){
        self$add_tlfmetadata(config)
      }
    },
    has_outputs = function(pattern = NULL) {
      outputs = ls(self$tlfs, all.names = TRUE)
      if (!is.null(pattern)) {
        assert_string(pattern)
        outputs = outputs[grepl(pattern, outputs)]
      }
      outputs
    },
    get = function(output, ...) {
      assert_string(output, min.chars = 1L)
      # return output
      obj = get0(output, envir = self$tlfs, inherits = FALSE, ifnotfound = NULL)
      return(obj)
    },
    mget = function(outputs, ...) {
      assert_character(outputs, min.chars = 1L, any.missing = FALSE)
      # return
      set_names(lapply(outputs, self$get, ...), outputs)
    },
    # read yaml configuration
    add_tlfmetadata = function(config) {
      assert_string(config, min.chars = 1L)
      assert(check_file_exists(config))

@@

    generator = function(VariableProc, tlfmetadata = self$tlfmetadata){

@@

      invisible(self)
    },
    remove = function(keys) {
      i = wf(!self$has(keys))
      rm(list = keys, envir = self$items)
      invisible(self)
    }
  )
)
```

initialize a new instance of `TLFProc`, add tlfmetadata `mock.yaml`
```{r}
tlfp <- TLFProc$new()

@@

<details><summary  class = "block2"> <font color="blue"> mock.yaml </font> __(click)__ </summary>

>>>>

---
title: 'Draft TLF generator'
date: '2023-04-25'
author: Qing Zou
categories: ['R', 'R6class', 'TLF']
description: 'Draft the TLF generator tool'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

@@

```{r rcell15}
#| echo: true
#| eval: true
#' @title AdsModel Class
#'
#'
#' @description
#' This is the  base class for ADaM dataset, to manage the CDISC attributes on dataset level
#'
#' AdsModel are build around the three following key parts:
#'
#' * Methods
#'
#'

AdsModel <- R6::R6Class(
    "AdsModel",
    public = list(
        initialize = function(...){
            df_lst <-  tibble::lst(...)
            if(is.character(check_list(df_lst, min.len = 1))){
                # message("No datasets passed")
            }else{
                df_lst <- purrr::map2(df_lst, names(df_lst), ~utils$add_model_name(.x, .y))
                private$.dataset <- private$.dataset |> append(df_lst)
            }

        },
        #' retrieve datasets from AdsModel by names
        get_dataset = function(domain){
            if(missing(domain)){
                return(private$.dataset)
            }else{
                checkmate::check_character(domain,
                                           min.chars = 1L,
                                           any.missing = FALSE)
                if(length(domain) == 1){
                    return(purrr::keep_at(private$.dataset, domain)[[1]])
                }else{
                    return(purrr::keep_at(private$.dataset, domain))
                }

@@

        },
        #' load dataset from disk
        #'
        #' @param path the database directory
        #' @param domain the dataset names
        load_dataset = function(path = NULL, domains){

@@

#### Used functions

<details><summary  class = "block2"> <font color="blue"> add_model_name </font> __(click)__ </summary>
```{r}
#| echo: true
#| eval: false
#' add a Model name attribute to data.frame
add_model_name <- function(df, name){
    attr(df, 'Model') <-toupper(name)
    return(df)
}
```
</details>

#### Steps
```{r}
#| echo: true
#| eval: true
# test <- AdsModel$new(adsl)
adsl <- haven::read_xpt('../../data/adam/adsl.xpt')
test <- AdsModel$new()

@@

# retrieve data by 'dataset'
names(test$dataset)
```

@@

tmp <- test$get_dataset('adsl')
#the updated attribute 'Model'
attributes(tmp)$Model
```

@@

adtte <- haven::read_xpt('../../data/adam/adtte.xpt')

# add dataset to instance
test$add_dataset(adtte)
names(test$dataset)
tmp2 <- test$get_dataset('adtte')
attributes(tmp2)$Model
```

@@

VariableProc <- R6::R6Class(
  "VariableProc",
  inherit = AdsModel,
  public = list(
    # save the dictionary
    dictionary = list(),
    initialize = function(dictionary_path = NA){

@@

- the `VariableProc` inherit from `AdsModel`
- method `get_decode` set attributes format to variables

#### Used functions

<details><summary  class = "block2"> <font color="blue"> set_attr </font> __(click)__ </summary>
```{r}
#| echo: true
#| eval: false
#' add "format" attributes to the variables of passed datasets
#'
#' @param df
#' @param dictionary
#' @param .vars the variables used to be assisgned the attributes
set_attr <- function(df, dictionary, .vars){

    sub_dict <- purrr::keep_at(dictionary, colnames(df))

    for (var in names(sub_dict)) {
        var_N <- ifelse(paste0(var, 'N') %in% colnames(df), paste0(var, 'N'), 'NA')
        attr(df[[var]], "format") <- list("Value" =  names(sub_dict[[var]]),
                                          "Decode" = purrr::map_vec(sub_dict[[var]], 1),
                                          "Numeric variable" = var_N,
                                          "Ordinal" = names(sub_dict[[var_N]]))
    }

    return(df)
}

```
</details>

#### Steps

@@

# inherit the AdsModel method
attributes(adsl_tmp)$Model

# before setting the format
attributes(adsl_tmp$SEX)$format

@@

# set the format by get_decode()
test2$get_decode()

adsl_tmp2 <- test2$get_dataset('adsl')

# the format is set from dictionary.json
attributes(adsl_tmp2$SEX)$format

@@

```{r}
TLFProc <- R6::R6Class(
  "TLFProc",
  public = list(

    #' Stores the object of the outputs
    tlfs = NULL,

    #' Store the metadata
    tlfmetadata = NULL,

    #' create a new environment for TLFProc
    initialize = function(config = NULL) {
      self$tlfs = new.env(parent = emptyenv())
      if(!is.null(config)){
        self$add_tlfmetadata(config)
      }
    },
    #' return avaialable outputs name
    has_outputs = function(pattern = NULL) {
      outputs = ls(self$tlfs, all.names = TRUE)
      if (!is.null(pattern)) {
        assert_string(pattern)
        outputs = outputs[grepl(pattern, outputs)]
      }
      outputs
    },

    #' get output by name
    get = function(output, ...) {
      assert_string(output, min.chars = 1L)
      # return output
      obj = get0(output, envir = self$tlfs, inherits = FALSE, ifnotfound = NULL)
      return(obj)
    },

    #' get a list of outputs
    mget = function(outputs, ...) {
      assert_character(outputs, min.chars = 1L, any.missing = FALSE)
      # return
      set_names(lapply(outputs, self$get, ...), outputs)
    },
    # read yaml configuration
    add_tlfmetadata = function(config) {
      assert_string(config, min.chars = 1L)
      assert(check_file_exists(config))

@@

    #' Generate the TLFs
    #' @param VariableProc the instance VariableProc with decoded dataset
    #' @param tlmetadata  the passed metadata
    generator = function(VariableProc, tlfmetadata = self$tlfmetadata){

@@

      invisible(self)
    },
    remove = function(keys) {
      i = wf(!self$has(keys))
      rm(list = keys, envir = self$items)
      invisible(self)
    }
  )
)
```
#### Used functions

<details><summary  class = "block2"> <font color="blue"> parse_mtdata </font> __(click)__ </summary>
```{r}
#| echo: true
#| eval: false
parse_mtdata <- function(metadata){

    # check the structure of metadata

    #------------------------------#
    #------------------------------#

    #check the results

    return(list(variables = names(metadata$columns),
                string_variable = names(purrr::keep(metadata$columns, ~.x$type == "string")),
                domains = purrr::map_vec(metadata$columns, ~purrr::pluck(.x, 'domain'))  |>  unique(),
                header_variable = purrr::pluck(metadata, "header")$variable,
                header_order = purrr::pluck(metadata, "header")$order,
                population_variable = purrr::pluck(metadata, "population")$variable)
    )
}
```
</details>

<details><summary  class = "block2"> <font color="blue"> demo_t_generator </font> __(click)__ </summary>
```{r}
#| echo: true
#| eval: false

demo_t_generator <- function(dataset, columns, columns_label, string_variable, header_variable, header_order, population_variable){

    names(header_variable) <- NULL

    adsl <- dataset %>%
        select(USUBJID, all_of(population_variable), all_of(header_variable), all_of(header_order), all_of(columns)) %>%
        filter(!!as.name(population_variable) == "Y")
    # %>%
    #   mutate(across(string_variable, ~ifelse(.=="", "Missing_value_impute", as.character(.))))

    adsl_fact <- adsl %>%
        mutate(across(all_of(string_variable),
                      ~factor(attributes(cur_data()[[cur_column()]])$format$Decode[.x],
                              levels = attributes(cur_data()[[cur_column()]])$format$Decode)))

    if(isTRUE(is.na(header_order))){
        adsl_fact <- adsl_fact %>%
            mutate(across(all_of(header_variable),
                          ~factor(attributes(cur_data()[[cur_column()]])$format$Decode[.x],
                                  levels = attributes(cur_data()[[cur_column()]])$format$Decode)))
    }else{
        adsl_fact <- df_order_header(adsl_fact, c(header_variable, header_order))
    }

    col_counts <- table(adsl[[header_variable]]) %>% as.vector()
    col_counts <- c(col_counts, sum(col_counts))

    calculated_t <- basic_table() %>%
        split_cols_by(header_variable) %>%
        add_overall_col("All") %>%
        summarize_vars(
            vars = columns,
            var_labels = columns_label,
            dispfmt = c("npct"),
            decimal = 1,
            denom = c("col"),
            unique_row = TRUE
        ) %>%
        rtables::build_table(adsl_fact, col_counts = col_counts)

    return(calculated_t)
}

```
</details>

#### Steps
- initialize a new instance of `TLFProc`, add tlfmetadata `mock.yaml`
```{r}
tlfp <- TLFProc$new()

@@

<details><summary  class = "block2"> <font color="blue"> demo_t_generator </font> __(click)__ </summary>
